MSTK: Mesh Toolkit, v 1.7 - DRAFT

### 2010-02-25 docs/MSTK.tex
Fixed error in description of MF_Coords
Updated!

\title{MSTK: Mesh Toolkit, v 1.3 - DRAFT}

\author{Rao V. Garimella \\
  Mathematical Modeling and Analysis (T-7), Theoretical Division, \\
  Los Alamos National Laboratory, Los Alamos, NM, USA. \\ E-mail:
  rao@lanl.gov}

@@

MSTK or Mesh Toolkit is a mesh framework that allows users to
represent, manipulate and query unstructured 3D arbitrary topology
meshes in a general manner without the need to code their own data
structures. MSTK is a flexible framework in that it allows (or will
eventually allow) a wide variety of underlying representations for the
mesh while maintaining a common interface. It will allow users to
choose from different mesh representations either at initialization or
during the program execution so that the optimal data structures are
used for the particular algorithm. The interaction of users and
applications with MSTK is through a functional interface that acts as
though the mesh always contains vertices, edges, faces and regions and
maintains connectivity between all these entities.

@@

Garimella, R. ``Mesh Data Structure Selection for Mesh Generation and
FEA Applications,'' \textit{International Journal of Numerical Methods
  in Engineering}, v55 n4, pp. 441-478, 2002.

In the following sections, the data types of MSTK will be described
followed by a description of the functional interface. The MSTK file
format will be described in the last section.

@@

\begin{description}
\item[]\textbf{\textit{List\_ptr}}: Handle to a List object.

\item[]\textbf{\textit{Mesh\_ptr}}: Handle to a Mesh object.

\item[]\textbf{\textit{MVertex\_ptr}}: Handle to a Mesh Vertex object (Topological
Dimension 0).

\item[]\textbf{\textit{MEdge\_ptr}}: Handle to a Mesh Edge object (Topological Dimension 1).

\item[]\textbf{\textit{MFace\_ptr}}: Handle to a Mesh Face object (Topological Dimension 2).

\item[]\textbf{\textit{MRegion\_ptr}}: Handle to Mesh Region object (Topological Dimension 3).

\item[]\textbf{\textit{MEntity\_ptr}}: Handle to a generic Mesh Entity
  object. Any of the above types of entities can be cast as
  \textit{MEntity\_ptr}.

\item[]\textbf{\textit{GModel\_ptr}}: Handle to a Geometric Model object.
\item[]\textbf{\textit{GEntity\_ptr}}: Handle to a Geometric Entity object.

\item[]\textbf{\textit{MAttrib\_ptr}}: Handle to a mesh attribute.

\item[]\textbf{\textit{RepType}}: Enumerated type describing the type
  of mesh representation.  Can be UNKNOWN\_REP, F1, F4, R1, R2, R4.
  See Appendix~\ref{app:reptypes} for schematics of these
  representations. \textit{Currently only representation types F1 and
    F4 are supported}.

\item[]\textbf{\textit{MType}}: Enumerated type for mesh entity type. Can be MDELETED, MVERTEX, MEDGE, MFACE, MREGION, MUNKNOWNTYPE, MALLTYPE.

\item[]\textbf{\textit{MFType}}: Enumerated type for mesh face type.
Can be FDELETED, FUNKNOWN, TRI, QUAD, POLYGON.

\item[]\textbf{\textit{MRType}}: Enumerated type for mesh region type.
Can be RDELETED, RUNKNOWN, TET, PYRAMID, PRISM, HEX, POLYHED.

\item[]\textbf{\textit{AttType}}: Enumerated type for attribute data. Can be INT, DOUBLE or POINTER
\end{description}

@@

Lists of entities in MSTK are returned as sets of type
\textit{List\_ptr}. The following are the set operations available in
MSTK:

\begin{description}
\item[]\textbf{\textit{List\_ptr} List\_New(\textit{int} inisize):} Create a
new set with an initial size, \textit{inisize}. If \textit{inisize} is
0, the initial size is set to be 10.

\item[]\textbf{\textit{void} List\_Delete(\textit{List\_ptr} l):} Delete a set.

\item[]\textbf{\textit{List\_ptr} List\_Compress(\textit{List\_ptr}
    l):} Compress a set.  Doing this while an algorithm is iterating
  through the set can currently cause problems!! Calling
  \textbf{List\_Compress} could change the pointer for the set due to
  reallocation.

\item[]\textbf{\textit{List\_ptr} List\_Copy(\textit{List\_ptr} l):} Return a
copy of a set.

\item[]\textbf{\textit{List\_ptr} List\_Add(\textit{List\_ptr} l, \textit{void
  *}entry):} Add an entry to the set. The entry is appended to the end
of the set.

\item[]\textbf{\textit{List\_ptr} List\_ChknAdd(\textit{List\_ptr} l,
\textit{void *}entry):} Add an entry to a set only if it is not already
in the set.

\item[]\textbf{\textit{int} List\_Rem(\textit{List\_ptr} l, \textit{void
  *}entry):} Remove an entry from the set. Returns 1 if successful, 0
otherwise.

\item[]\textbf{\textit{int} List\_Remi(\textit{List\_ptr} l, \textit{int} i):}
Remove the i'th valid entry in the set. Returns 1 if successful, 0
otherwise.

\item[]\textbf{\textit{int} List\_Replace(\textit{List\_ptr} l, \textit{void
  *}entry, \textit{void *}nuentry):} Replace 'entry' with 'nuentry' in
set. Returns 1 if successful, 0 otherwise.

\item[]\textbf{\textit{int} List\_Replacei(\textit{List\_ptr} l, \textit{int}
i, void *nuentry):} Replace the i'th valid entry in the set with
'nuentry'. Returns 1 if successful, 0 otherwise.

\item[]\textbf{\textit{int} List\_Contains(\textit{List\_ptr} l, void *entry):}
Returns 1 if set contains the entry, 0 otherwise.

\item[]\textbf{\textit{int} List\_Locate(\textit{List\_ptr} l, void *entry):}
Returns the positional index of the entry in the set. Returns -1 if
the set does not contain the entry.

\item[]\textbf{\textit{void *}List\_Entry(\textit{List\_ptr} l, \textit{int}
i):} Return the i'th valid entry in the set.  Returns a NULL pointer if
the i'th valid entry could not be found.

\item[]\textbf{\textit{void *}List\_Next\_Entry(\textit{List\_ptr} l,
\textit{int} *i):} Return the next valid entry in the set. This routine
works like an iterator. To start iterating through the set, set the
iteration index i=0 and call the routine to get the first entry in the
set. Subsequent calls to the routine will iterate through the entries
in the set. The routine will return a NULL to indicate that the end of
the set is reached.

The value of the iteration index i will be modified by the routine on
each call to indicate where in the set it is. This value should not be
modified externally while iterating through the set. Also, no specific
meaning should be derived from the iteration index by other applications
since the internal implementation and interpretation of the index may
change at any time.

\item[]\textbf{\textit{int} List\_Num\_Entries(\textit{List\_ptr} l):} Return
the number of entries in a set
\end{description}

@@

\begin{description}
\item[]\textbf{\textit{Mesh\_ptr} MESH\_New(\textit{RepType} type):}
Initialize a new mesh object with the given representation type which
can be F1, F2, F3, F4, F5, F6, R1, R2, R3, R4. Not all of these types
are implemented. If the representation type is not known at the
present time (e.g. before reading the mesh from a file), the
representation type of UNKNOWN\_REP can be specified.  Note that this
only initializes a mesh object, it does not create or generate a mesh
which is the work of high level mesh generation routines.

\item[]\textbf{\textit{int} MESH\_InitFromFile(\textit{Mesh\_ptr} mesh,
const char *filename):} Initialize or read a mesh from file into the
given mesh object. Returns 1 if successful, 0 otherwise.

\item[]\textbf{\textit{int} MESH\_ImportFromFile(\textit{Mesh\_ptr}
    mesh, const char *filename, const char *format):} Import a mesh
  data from an external file format and construct MSTK mesh. Currently
  the only format supported is the
  GMV\footnote{\texttt{http://www-xdiv.lanl.gov/XCM/gmv/GMVHome.html}}
  file format. The routine imports as many attributes as it can from
  the input file. It also uses special attributes or keywords as data
  about element and node classification.  For GMV files, the routine
  uses the ``material'' data to indicate region or face classification
  depending on the dimensionality of the mesh. If ``material'' data is
  not specified, it uses the ``itetclr'' attribute to assign region or
  face classification. It also uses the ``icr'' keyword describing the
  number of constraints on a node to interpret if the node is
  classified on a model region, model face, model edge or a model
  vertex. The ``itetclr'' and ``icr'' keywords are usually present in
  meshes generated by
  LAGRIT\footnote{\texttt{http://lagrit.lanl.gov'}}.

\item[]\textbf{\textit{int} MESH\_BuildClassfn(\textit{Mesh\_ptr}
    mesh):} Build classification information for mesh entities if only
  partial information is present. In other words, if the only data
  known is the IDs of geometric model entities (``material regions'')
  of the highest level mesh entities (faces or regions), then this
  procedure will build information about the type of geometric model
  entities that the lower dimension entities are on. Faces will be
  classified as being on a model face (external or interior) or inside
  model region. Edges will be classified as being on a model edge,
  model face or in a model region. Vertices will be classified as
  being on a model vertex, model edge, model face or in a model
  region. If no classification data is associated with even the
  highest level entities, the procedure will assume that all the
  highest level entities are classified on one model entitiy of that
  dimension. If partial information is available for lower order
  entities, this routine will not destroy that information. In the
  future, the procedure will be more sophisticated, incorporating
  features such as classifying mesh edges as being on model edges if
  the dihedral angle between the boundary faces connected to it is
  very large.

\item[]\textbf{\textit{int} MESH\_DelInterior(\textit{Mesh\_ptr}
    mesh):} Delete the interior of a mesh and retain only its
    boundaries (including interior boundaries). For solid meshes, this
    results in a surface mesh (i.e., all mesh regions, and  mesh
    faces, edges and vertices classified on model regions are deleted). For
    surface meshes, this results in curve mesh. For an edge mesh, only
    the end vertices are retained (unlikely to be used this
    way). Undefined for a vertex mesh.

\item[]\textbf{\textit{void} MESH\_WriteToFile(\textit{Mesh\_ptr}
    mesh, const char *filename):} Save a mesh to a filename. The file
  is created if it does not exists. It is recommended that the .mstk
  extension be used for MSTK mesh files.  However, there is no such
  requirement.

%% \item[]\textbf{\textit{void} MESH\_ExportToFile(\textit{Mesh\_ptr}
%%     mesh, const char *filname, const char *format, int natt, char
%%     **attnames):} Export a mesh to an external file format. Currently,
%%   the formats supported are the GMV file (format string: ``gmv'') and
%%   Data Explorer\footnote{\texttt{http://opendx.org}} (format string:
%%   ``dx''). Only integer and double attributes of the mesh are exported
%%   to the files.

\item[]\textbf{\textit{void} MESH\_ExportToFile(\textit{Mesh\_ptr}
    mesh, const char *filname, const char *format, int natt, char
    **attnames):} Export a mesh to an external file format. Currently,
  the only format supported is the GMV file format (format string:
  ``gmv''). Only integer and double attributes of the mesh are
  exported to the files.

@@

\item[]\textbf{\textit{GModel\_ptr} MESH\_GModel(\textit{Mesh\_ptr} mesh):}
Return a handle to the underlying geometric model. If there is no
geometric model associated with the mesh, NULL pointer is returned.

\item[]\textbf{\textit{RepType} MESH\_RepType(\textit{Mesh\_ptr} mesh):}
Representation type currently being used by the mesh.

@@

\item[]\textbf{\textit{int} MESH\_Num\_Vertices(\textit{Mesh\_ptr} mesh):}
Number of vertices in the mesh.

\item[]\textbf{\textit{int} MESH\_Num\_Edges(\textit{Mesh\_ptr} mesh):}
Number of edges in the mesh. For reduced representations, this routine
returns 0 since it is impractically expensive to count the number of
edges when they do not explicitly exist. Applications must find a way
to avoid using this routine for reduced representations.

\item[]\textbf{\textit{int} MESH\_Num\_Faces(\textit{Mesh\_ptr} mesh):}
Number of faces in the mesh. For reduced representations R1 or R2,
this routine counts only the faces that are explicitly represented
i.e. faces not connected to any mesh region. Therefore, a value of 0
will be returned for the number of faces of a tetrahedral mesh with
representation R1 or R2 but the correct number will be reported for a
tetrahedral mesh in other representations. Also, the correct number
will be reported for the number of faces in a surface mesh in
representation R1 or R2.  Therefore, this routine must be used
carefully.

\item[]\textbf{\textit{int} MESH\_Num\_Regions(\textit{Mesh\_ptr} mesh):}
Number of regions in the mesh.

@@

\item[]\textbf{\textit{MVertex\_ptr} MESH\_Vertex(\textit{Mesh\_ptr} mesh,
\textit{int} i):} Return the i'th vertex in the mesh. Returns NULL if i
$<$ 0 or i $>$ number of mesh vertices.

>>>>

\title{MSTK: Mesh Toolkit, v 1.7 - DRAFT}

\author{Rao V. Garimella \\
  T-5, Theoretical Division, \\
  Los Alamos National Laboratory, Los Alamos, NM, USA. \\ E-mail:
  rao@lanl.gov}

@@

\subsection{What is MSTK?}
MSTK or Mesh Toolkit is a mesh framework that allows users to
represent, manipulate and query unstructured 3D arbitrary topology
meshes in a general manner without the need to code their own data
structures.

\subsection{What MSTK is not?}

MSTK is not a mesh generator - it can be used to write more easily
than starting from scratch. Also, MSTK cannot answer computational
questions related to the mesh (for example, is this point in the given
mesh element?).

\subsection{Salient Features of MSTK}

MSTK is a flexible framework in that it allows a variety
of underlying representations for the mesh while maintaining a common
interface. It will allow users to choose from different mesh
representations either at initialization (implemented) or during the
program execution (not implemented) so that the optimal data
structures are used for the particular algorithm. The interaction of
users and applications with MSTK is through a functional interface
that acts as though the mesh always contains vertices, edges, faces
and regions and maintains connectivity between all these entities.

@@

Garimella, R. ``Mesh Data Structure Selection for Mesh Generation and
FEA Applications,'' {\em International Journal of Numerical Methods
  in Engineering}, v55 n4, pp. 441-478, 2002.

\subsection{Why should I use MSTK?}

MSTK offers a flexible infrastructure for representing and
manipulating meshes so that mesh-based application developers do not
have to deal with the complexitiies of managing their own data
structures. The functional interface of MSTK is designed to provide
easy access to mesh data and allow for easy-to-read but efficient code
to be written for mesh based applications. Accessing the mesh data
through the functional interface allows the high level application to
be unchanged even if the lower level data structures in MSTK
change. MSTK supports a wide array of element types and several
representations, all of which can take a new developer years to write
and make robust. New users of MSTK can typically start writing their
own codes to query and manipulate meshes within a few days of
familiarizing themselves with MSTK. Finally, use of MSTK allows easy
integration with other codes using MSTK as their mesh data framework.

\section{MSTK Concepts}
\subsection{Unstructured mesh representation}

Meshes are made up of topological entities of different dimensions. In
a ``traditional'' 3D finite element mesh, nodes are topological
entities called vertices and are topologically 0-dimensional entities,
while elements are topological entities called regions and are
topologically 3-dimensional entities. In general, meshes can be
described using vertices (0-dimensional entities), edges
(1-dimensional entities), faces (2-dimensional entities) and regions
(3-dimensional entities). MSTK has multiple ways of representing
meshes as shown in the below. For example, representation F1 has
entities of all dimensions up to the dimension of the mesh while
representation R1 has entities of the lowest and highest dimension
only. Currently, the application has to choose a particular
representation type when a mesh is created although it may choose
different representations for different meshes in the same program. In
the future, the code will be allowed to switch between different
representations depending on the needs of the particular algorithm
being executed at that time.

{\bf NOTE: CURRENTLY THE F1 REPRESENTATION IS THE MOST ROBUST
  IMPLEMENTATION SINCE IT IS MOST COMMONLY USED. THIS REPRESENTATION
  HAS GENERALLY BEEN STABLE FOR SEVERAL YEARS. THE REDUCED
  REPRESENTATION HAVE NOT BEEN AS HEAVILY DEBUGGED AND IT IS CURRENTLY
  BROKEN. IF YOU CHOOSE TO USE SOME OF THE OTHER REPRESENTATIONS
  PLEASE REPORT ANY PROBLEMS IMMEDIATELY TO \texttt{rao@lanl.gov}.  }

Even when using reduced representations, MSTK can provide the full set
of topological adjacencies as if a full representation were being
used. When necessary, entities not represented explicitly in the mesh
are created on-the-fly and represented as volatile entities. Even
though it is possible to take an algorithm that is written for a full
representation and use it as is with a reduced representation,
application developers are urged to be aware of the costs of each
operation for different representations and use this information to
design algorithms optimized for each representation.

\subsection{Mesh entity classification}

The concept of mesh entities of various dimensions is actually derived
from field of B-rep geometric modeling in which a model is comprised
of a hierarchy of topological entities of different dimensions. {\em
  Classification} is the relationship of each mesh entity to the model
entity is represents the whole or a part of.

To clarify further, every mesh is a discrete representation of a
geometric model. This geometric model may exist in the analyst's mind,
as schematic on paper or as a full model in a geometric modeling
system. Encoding the relationship of the mesh to this geometric model
to the extent possible provides enormous algorithmic benefits to the
users of meshes.

Recognizing that one may have varying degrees of information about the
geometric model in different situations, MSTK allows mesh entities to
store different levels of classification information or no
classification information.

Every mesh entity can store the dimension of the model entity it is on
({\bf GEntDim}), the ID of the model entity it is on ({\bf GEntID})
and a pointer to the model entity it is on ({\bf GEntity}). The
following picture illustrates the classification of various entities
in a simple mesh.

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{figures/classfn}
  \end{center}
  \caption{Example mesh and model showing the classification of mesh entities on geometric model entities}
\end{figure}

Classification information is very useful to have for simplifying
algorithms in meshing and analysis applications. For example, using
classification information, one can trivially extract all nodes on the
boundary and apply a different smoothing algorithm than nodes in the
interior of the domain. An FEA code can apply boundary conditions on
nodes and faces more easily if it has classification information.

If the application developer has nothing but the material IDs of the
mesh elements, MSTK can build the classification information to the
best of its ability using topological and geometric
information. However, it must be noted that this is procedure is not
perfect and some classification information cannot be derived
unambiguously. So, wherever possible, users are encouraged to generate
classification information during mesh generation and supply it to
MSTK.

\subsection{MSTK Application Programming Interface (API)}

Even though MSTK is written in C, MSTK is designed in an object
oriented manner. Mesh entities, mesh attributes and entity lists are
considered to be objects which have their private data and a set of
operators to query and manipulate this data. Even though methods may
be devised to circumvent the data hiding in MSTK, this is highly
discouraged. If you think that a particular operator is not efficient
or you don'thave access to some mesh data, please contact the
developer for assistance.

An important feature of MSTK is its ability to provide the full set of
mesh operators regardless of the representation used. When using
reduced representations, entities not represented explicitly in the
mesh are returned as volatile objects. However, it is possible to use
most MSTK operators on such objects as well.

Currently, MSTK offers only a C API but C++ and Fortran APIs are
planned for the near future.

\subsection{Getting meshes in and out of MSTK}

There are a few ways of getting mesh data into MSTK. The first is to
prepare an MSTK file with the full connectivity of an F1
representation and to read it in using MESH\_InitFromFile. This is
typically a difficult task for many applications which only have
element-node data. In such cases, one can prepare an MSTK file in the
R1 file format (in which the node listing is followed by the
region-node connectivity) and read it in to any of the other
representations, say F1. The other possibility is to write a file in
the GMV format and use MESH\_ImportFromFile to import the mesh in.

MSTK can write meshes out to different formats including GMV, STL and
FLAG X3D. When compiled with the appropriate options and using a
partitioner from an external library, MSTK can also write out a
parallel X3D file.

@@

\begin{description}
\item[]{\bf {\em List\_ptr}}: Handle to a List object.

\item[]{\bf {\em Mesh\_ptr}}: Handle to a Mesh object.

\item[]{\bf {\em MVertex\_ptr}}: Handle to a Mesh Vertex object (Topological
Dimension 0).

\item[]{\bf {\em MEdge\_ptr}}: Handle to a Mesh Edge object (Topological Dimension 1).

\item[]{\bf {\em MFace\_ptr}}: Handle to a Mesh Face object (Topological Dimension 2).

\item[]{\bf {\em MRegion\_ptr}}: Handle to Mesh Region object (Topological Dimension 3).

\item[]{\bf {\em MEntity\_ptr}}: Handle to a generic Mesh Entity
  object. Any of the above types of entities can be cast as
  {\em MEntity\_ptr}.

\item[]{\bf {\em GModel\_ptr}}: Handle to a Geometric Model object.
\item[]{\bf {\em GEntity\_ptr}}: Handle to a Geometric Entity object.

\item[]{\bf {\em MAttrib\_ptr}}: Handle to a mesh attribute.

\item[]{\bf {\em RepType}}: Enumerated type describing the type
  of mesh representation.  Can be UNKNOWN\_REP, F1, F4, R1, R2, R4.
  See Appendix~\ref{app:reptypes} for schematics of these
  representations. {\em Currently only representation types F1 and
    F4 are supported}.

\item[]{\bf {\em MType}}: Enumerated type for mesh entity type. Can be MDELETED, MVERTEX, MEDGE, MFACE, MREGION, MUNKNOWNTYPE, MALLTYPE.

\item[]{\bf {\em MFType}}: Enumerated type for mesh face type.
Can be FDELETED, FUNKNOWN, TRI, QUAD, POLYGON.

\item[]{\bf {\em MRType}}: Enumerated type for mesh region type.
Can be RDELETED, RUNKNOWN, TET, PYRAMID, PRISM, HEX, POLYHED.

\item[]{\bf {\em AttType}}: Enumerated type for attribute data. Can be INT, DOUBLE or POINTER
\end{description}

@@

Unordered sets or lists of entities in MSTK are returned as type
{\em List\_ptr}. The following are the set operations available in
MSTK:

\begin{description}
\item[]{\bf {\em List\_ptr} List\_New({\em int} inisize):} Create a
new list with an initial size, {\em inisize}. If {\em inisize} is
0, the initial size is set to be 10.

\item[]{\bf {\em void} List\_Delete({\em List\_ptr} l):} Delete a list.

\item[]{\bf {\em List\_ptr} List\_Compress({\em List\_ptr}
    l):} Compress a list (delete all the null entries created when
  entities are removed from the list).  Doing this while an algorithm
  is iterating through the set can currently cause problems!! Calling
  {\bf List\_Compress} could change the pointer for the list due to
  reallocation.

\item[]{\bf {\em List\_ptr} List\_Copy({\em List\_ptr} l):} Return a
copy of a list.

\item[]{\bf {\em List\_ptr} List\_Add({\em List\_ptr} l, {\em void
  *}entry):} Add an entry to the list. The entry is strictly appended to the end
of the list.

\item[]{\bf {\em List\_ptr} List\_ChknAdd({\em List\_ptr} l,
{\em void *}entry):} Add an entry to a list only if it is not already
in the list.

\item[]{\bf {\em List\_ptr} List\_Insert({\em List\_ptr} l,
{\em void *}nuentry, {\em void *}b4entry):} Insert an entry into the list in the position before 'b4entry'.

\item[]{\bf {\em List\_ptr} List\_Inserti({\em List\_ptr} l,
    {\em void *}nuentry, {\em int} i):} Insert an entry into the
  list at the i'th valid position and push the entries previously at
  the i'th and later position back.

\item[]{\bf {\em int} List\_Rem({\em List\_ptr} l, {\em void
  *}entry):} Remove an entry from the list. Returns 1 if successful, 0
otherwise.

\item[]{\bf {\em int} List\_Remi({\em List\_ptr} l, {\em int} i):}
Remove the i'th valid entry in the list. Returns 1 if successful, 0
otherwise.

\item[]{\bf {\em int} List\_RemSorted({\em List\_ptr} l, {\em void}
    *entry, int (*entry2int)(void *)):} Remove an entry from a list
  that is sorted according to some correspondence between an entry and
  an integer value, e.g., {\bf MEnt\_ID(entry)} gives the ID of an
  entity. The mapping between the entry and the integer is given by a
  call to the routine {\bf entry2int}. The list could be one from which
  entities have previously been removed. The routine makes use of the
  sorted nature of the list to locate entries in $O(log(n))$ time
  where $n$ is the size of the list. An example call to this routine
  is

  \par {\bf fnd = List\_RemSorted(mregionslist,region,\&(MR\_ID));}

\item[]{\bf {\em int} List\_Replace({\em List\_ptr} l, {\em void
  *}entry, {\em void *}nuentry):} Replace 'entry' with 'nuentry' in
set. Returns 1 if successful, 0 otherwise.

\item[]{\bf {\em int} List\_Replacei({\em List\_ptr} l, {\em int}
i, void *nuentry):} Replace the i'th valid entry in the list with
'nuentry'. Returns 1 if successful, 0 otherwise.

\item[]{\bf {\em int} List\_Contains({\em List\_ptr} l, void *entry):}
Returns 1 if list contains the entry, 0 otherwise.

\item[]{\bf {\em int} List\_Locate({\em List\_ptr} l, void *entry):}
Returns the positional index of the entry in the list. Returns -1 if
the list does not contain the entry.

\item[]{\bf {\em void *}List\_Entry({\em List\_ptr} l, {\em int}
i):} Return the i'th valid entry in the list.  Returns a NULL pointer if
the i'th valid entry could not be found.

\item[]{\bf {\em void *}List\_Next\_Entry({\em List\_ptr} l,
{\em int} *i):} Return the next valid entry in the list. This routine
works like an iterator. To start iterating through the list, set the
iteration index i=0 and call the routine to get the first entry in the
set. Subsequent calls to the routine will iterate through the entries
in the list. The routine will return a NULL to indicate that the end of
the set is reached.

The value of the iteration index i will be modified by the routine on
each call to indicate where in the list it is. This value should not be
modified externally while iterating through the list. Also, no specific
meaning should be derived from the iteration index by other applications
since the internal implementation and interpretation of the index may
change at any time.

Finally, there may be unexpected consequences if entries are removed
from the list while an iterator is iterating through it using {\bf
  List\_Next\_Entry}.

\item[]{\bf {\em int} List\_Num\_Entries({\em List\_ptr} l):} Return
the number of entries in a list.
\end{description}

@@

\begin{description}
\item[]{\bf {\em Mesh\_ptr} MESH\_New({\em RepType} type):}
  Initialize a new mesh object with the given representation type
  which can be F1, F2, F3, F4, F5, F6, R1, R2, R3, R4. Of these types
  F1, F4, R1, R2 and R4 are implemented. If the representation type is
  not known at the present time (e.g. before reading the mesh from a
  file), the representation type of UNKNOWN\_REP can be specified.
  Note that this only initializes a mesh object, it does not create or
  generate a mesh which is the work of high level mesh generation
  routines.

\item[]{\bf {\em int} MESH\_InitFromFile({\em Mesh\_ptr}
    mesh, const char *filename):} Initialize or read a mesh from file
  into the given mesh object. Returns 1 if successful, 0 otherwise. It
  is possible to have a an MSTK file in the R1, R2 and R4 format into
  a mesh initialized as type F1 or F4.

\item[]{\bf {\em int} MESH\_InitFromGenDesc({\em Mesh\_ptr} mesh, {\em
      int} nv, {\em double} (*xyz)[3], {\em int} *nfv, {\em int}
    **fvids, {\em int} nr, {\em int} *nrv, {\em int} **rvids, {\em
      int} *nrf, {\em int} ***rfvtemplate):} Initialize a mesh from a
  minimal description of the mesh passed into the routine. In the
  routine, {\bf nv} is the number of vertices or nodes and {\bf xyz}
  is the array of node coordinates. If the mesh is a surface mesh,
  {\bf nf} is the number of mesh faces, {\bf nfv} gives the number of
  vertices for each face, {\bf fvids} gives the array indices of the
  face vertices in ccw manner (starting from 0, not 1). If the mesh is
  a solid mesh, {\bf nr} specifies the number of solid elements or
  regions. If the mesh has only standard element types (tets,
  pyramids, triangular prisms and hexes), then {\bf nrv} indicates the
  number of vertices of each region and {\bf rvids} gives the array
  indices of the region vertices. If, on the other hand, the mesh has
  polyhedral elements, regions have to be described in terms of faces
  which are in turn described in terms of vertices. Therefore, {\bf
    nrf} indicates the number of faces for each region and {\bf
    rfvtemplate} gives the array indices of the vertices for each face
  of the region.

\item[]{\bf {\em int} MESH\_ImportFromFile({\em Mesh\_ptr}
    mesh, const char *filename, const char *format):} Import a mesh
  data from an external file format and construct MSTK mesh. Currently
  the only format supported is the
  GMV\footnote{\texttt{http://www-xdiv.lanl.gov/XCM/gmv/GMVHome.html}}
  file format. The routine imports as many attributes as it can from
  the input file. It also uses special attributes or keywords as data
  about element and node classification.  For GMV files, the routine
  uses the ``material'' data to indicate region or face classification
  depending on the dimensionality of the mesh. If ``material'' data is
  not specified, it uses the ``itetclr'' attribute to assign region or
  face classification. It also uses the ``icr'' keyword describing the
  number of constraints on a node to interpret if the node is
  classified on a model region, model face, model edge or a model
  vertex. The ``itetclr'' and ``icr'' keywords are usually present in
  meshes generated by
  LAGRIT\footnote{\texttt{http://lagrit.lanl.gov'}}.

\item[]{\bf {\em int} MESH\_BuildClassfn({\em Mesh\_ptr} mesh):} Build
  classification information for mesh entities if only partial
  information is present. In other words, if the only data known is
  the IDs of geometric model entities (``material regions'') of the
  highest level mesh entities (faces or regions), then this procedure
  will build information about the type of geometric model entities
  that the lower dimension entities are on. Faces will be classified
  as being on a model face (external or interior) or inside model
  region. Edges will be classified as being on a model edge, model
  face or in a model region. Vertices will be classified as being on a
  model vertex, model edge, model face or in a model region. If no
  classification data is associated with even the highest level
  entities, the procedure will assume that all the highest level
  entities are classified on one model entitiy of that dimension. If
  partial information is available for lower order entities, this
  routine will not destroy that information. The procedure also tries
  to detect mesh edges that should be classified on model edges based
  on the fact that the dihedral angle between the boundary faces
  connected to the edge is smaller than some tolerance.

\item[]{\bf {\em int} MESH\_DelInterior({\em Mesh\_ptr}
    mesh):} Delete the interior of a mesh and retain only its
    boundaries (including interior boundaries). For solid meshes, this
    results in a surface mesh (i.e., all mesh regions, and  mesh
    faces, edges and vertices classified on model regions are deleted). For
    surface meshes, this results in curve mesh. For an edge mesh, only
    the end vertices are retained (unlikely to be used this
    way). Undefined for a vertex mesh.

\item[]{\bf {\em void} MESH\_WriteToFile({\em Mesh\_ptr}
    mesh, const char *filename):} Save a mesh to a filename. The file
  is created if it does not exists. It is recommended that the .mstk
  extension be used for MSTK mesh files.  However, there is no such
  requirement.

%% \item[]{\bf {\em void} MESH\_ExportToFile({\em Mesh\_ptr}
%%     mesh, const char *filname, const char *format, int natt, char
%%     **attnames):} Export a mesh to an external file format. Currently,
%%   the formats supported are the GMV file (format string: ``gmv'') and
%%   Data Explorer\footnote{\texttt{http://opendx.org}} (format string:
%%   ``dx''). Only integer and double attributes of the mesh are exported
%%   to the files.

\item[]{\bf {\em void} MESH\_ExportToFile({\em Mesh\_ptr} mesh, const
    char *filname, const char *format, int natt, char **attnames):}
  Export a mesh to an external file format. Currently, the only
  formats supported is the GMV file format (format string: ``gmv'')
  and X3D (format string: ``flag''). Only integer and double
  attributes of the mesh are exported to GMV files and no attributes
  are exported to X3D files.

\item[]{\bf {\em void} MESH\_ExportToGMV({\em Mesh\_ptr} mesh, const
    char *filname, const char *format, int natt, char **attnames):} Export a mesh to a GMV file format.

\item[]{\bf {\em void} MESH\_ExportToFLAGX3D({\em Mesh\_ptr} mesh, const
    char *filname, const char *format, int natt, char **attnames):} Export a mesh to the FLAG X3D format.

\item[]{\bf {\em void} MESH\_ExportToFLAGX3D\_Par({\em Mesh\_ptr} mesh,
    const char *filname, const char *format, int natt, char
    **attnames, int *procids):} Export a mesh to the FLAG X3D format
  given a partitioned mesh. The owning processor for each element is
  given in the array {\bf procids}.

\item[]{\bf {\em void} MESH\_ExportToSTL({\em Mesh\_ptr} mesh, const
    char *filname):} Export a mesh to the STL format (SURFACE MESHES ONLY!).

\item[]{\bf {\em void} MESH\_Tet2Hex({\em Mesh\_ptr} tetmesh, {\em
      Mesh\_ptr} *hexmesh):} Convert a tet mesh to a hex mesh by
  splitting the tets - the quality of the resulting hex mesh is
  usually not very good.

\item[]{\bf {\em void} MESH\_Renumber({\em Mesh\_ptr} mesh):} Renumber
  the entities of a mesh to avoid gaps in entity IDs. Note that in the
  current implementation, renumbering mesh entities can make removal
  of and searching for mesh entities much slower, so this must be
  avoided as much as possible.

\item[]{\bf {\em int} MESH\_PartitionWithMetis({\em Mesh\_ptr} mesh,
    {\em int} nparts, {\em int} **part):} Partition a mesh with the
  METIS libraries. 'part' is an array that contains the partition
  number for each mesh face (surface meshes) or mesh region (volume
  meshes).

  \par {\bf NOTE THAT MSTK HAS TO HAVE BEEN COMPILED USING THE COMMAND
    'make PAR=1' AND THE CALLING APPLICATION MUST LINK WITH THE METIS
    LIBRARY.}

@@

\item[]{\bf {\em GModel\_ptr} MESH\_GModel({\em Mesh\_ptr} mesh):}
Return a handle to the underlying geometric model. If there is no
geometric model associated with the mesh, NULL pointer is returned.

\item[]{\bf {\em RepType} MESH\_RepType({\em Mesh\_ptr} mesh):}
Representation type currently being used by the mesh.

\item[]{\bf {\em char } *MESH\_RepType({\em Mesh\_ptr} mesh):}
  Representation type currently being used by the mesh returned as a
  2-character string.

@@

\item[]{\bf {\em int} MESH\_Num\_Vertices({\em Mesh\_ptr} mesh):}
Number of vertices in the mesh.

\item[]{\bf {\em int} MESH\_Num\_Edges({\em Mesh\_ptr} mesh):}
Number of edges in the mesh. For reduced representations, this routine
returns 0 since it is impractically expensive to count the number of
edges when they do not explicitly exist. Applications must find a way
to avoid using this routine for reduced representations.

\item[]{\bf {\em int} MESH\_Num\_Faces({\em Mesh\_ptr} mesh):}
Number of faces in the mesh. For reduced representations R1 or R2,
this routine counts only the faces that are explicitly represented
i.e. faces not connected to any mesh region. Therefore, a value of 0
will be returned for the number of faces of a tetrahedral mesh with
representation R1 or R2 but the correct number will be reported for a
tetrahedral mesh in other representations. Also, the correct number
will be reported for the number of faces in a surface mesh in
representation R1 or R2.  Therefore, this routine must be used
carefully.

\item[]{\bf {\em int} MESH\_Num\_Regions({\em Mesh\_ptr} mesh):}
Number of regions in the mesh.

@@

\item[]{\bf {\em MVertex\_ptr} MESH\_Vertex({\em Mesh\_ptr} mesh,
{\em int} i):} Return the i'th vertex in the mesh. Returns NULL if i
$<$ 0 or i $>$ number of mesh vertices.

\item[]{\bf {\em MEdge\_ptr} MESH\_Edge({\em Mesh\_ptr} mesh,
{\em int} i):} Return the i'th edge in the mesh. Returns NULL if i $<$
0 or i $>$ number of mesh edges. Returns NULL for reduced
representations.
